---
  title: "R package GSMC: a simulation-free group sequential design with max-combo tests in the presence of non-proportional hazards"
  author: "[Lili Wang](https://lilywang.info)"
  date: "`r Sys.Date()`"
  output: github_document
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(collapse = TRUE,
                     comment = "#>",
                     fig.width=12, fig.height=8,
                     fig.path = "man/figures/README-")

```

## Purpose

This R package is to implement the proposed [simulation-free method for group sequential design with maxcombo tests](https://arxiv.org/abs/1911.05684 "Hey, direct me to the tedious methodology paper!")(GS-MC). The goal here is to improve the detection power of the weighted log-rank tests (WLRTs) when the non-proportional hazards are present. Moreover, the method is simulation-free, and allows multiple checks (interims) before ending the study. 

This R package is currently dependent on other R packages including [mvtnorm][3] and my another package on github [IAfrac][1]. I hope that this package will be improved gradually please let us know if you spot anything unclear or suspicious to you. 

For details about the calculation of the information and covariance using estimation (data-driven) and prediction methods please find the README file of [IAfrac][1]. 



## Preparation

To install and use this R package from Github, you will need to install another R package "devtools" and two packages on Github: [nphsim][2], [IAfrac][1]. Please uncomment the codes to install them. 

```{r installation, results = "hide",message=FALSE}
# install.packages("devtools")
# library(devtools)

#for information and covariance calculation; sample size computation using Hasegawa proposal
# install_github("lilywang1988/IAfrac")
library(IAfrac) 

#Generate data following nonproportional hazards
# install_github("keaven/nphsim")
library(nphsim) 

#for information and covariance calculation; sample size computation using Hasegawa proposal
# install_github("lilywang1988/GSMC")
library(GSMC) 

```



## Vignette 1: sample size caulcation and plots for GS-MC ($d$ or $n$ vs $\epsilon$)
### Caveats: 
- Make sure you install and activate all the packages: [IAfrac][1] and [nphsim][2]
- Here I only conduct prediction based on the exact survival distributions, in Vignett2, I will introduce the stochastic prediction and estimation method
- Two (M=2) tests are in the combo: $G_{0,0}(t)$ and $G_{0,1}(t)$
- Two (K=2) stages are included: one interim stage and one final stage
- The codes I used to prepare Figure 2 in the [methodology paper](https://arxiv.org/abs/1911.05684 "Hey, direct me to the tedious methodology paper!")

```{r vignette1}
#rm(list=ls())
#install.packages("ggplot2")
library(ggplot2)
fixed_death_p<-0.6
p<-1/2
tau_star<-21 # a safe technical end
tau<-18
R <- 14 # accrual period
omega <- (tau-R) # total follow-up time after the end of the acrrual period
eps.ls <- seq(0,3.5,0.05) # the vector of eps (delayed time) you would like to test
lambda <- log(2)/6 # median survival is 6 months
(theta <- 0.6) # the hazard ratio after the change point eps is 0.6
lambda.trt <- lambda*theta # treatment effect after the change point eps
alpha <- 0.025 # type I error
beta <- 0.1 # type II error

b <- 30 # number of subintervals at each time point --> for the use of sample.size_FH in IAfrac

# Define the Fleming-Harrington class of WLRTs
rho <- 0 
gamma <- 1

# Obtain the boundaries under for regular single variable tests (e.g. WLRT, SLRT)
x <- gsDesign(k=2, test.type=1, timing=0.6, sfu="OF", alpha=alpha, beta=beta,delta=-log(theta))
(z <- x$upper$bound)

# Preparing variables to store the results
n_MCGS<-d_MCGS<-n_FH_00_ls<-n_FH_01_ls<-d_FH_00_ls<-d_FH_01_ls<-NULL

ll<-0 # index for eps.ls
len=length(eps.ls)

pt<- proc.time() # timer
for(eps in eps.ls){
  ll=ll+1
  # Below is for the sample size calculation under H0
  size_FH_00 <- sample.size_FH(eps,p,b,tau,omega,lambda,lambda.trt,0, 0,alpha,beta)
  size_FH_01<-sample.size_FH(eps,p,b,tau,omega,lambda,lambda.trt,0, 1,alpha,beta)
  sum_D<-size_FH_00$sum_D # should be identical to size_FH_01$sum_D
  
  size_range<-c(max(min(size_FH_00$n,size_FH_01$n)-50,10),max(max(size_FH_00$n,size_FH_01$n+50),20))  # introduce 50 to allow the sample size to be slightly lower or higher than the boundaries defined by the two test statistics; alternatively, one may use [10, 500] if computation time is not a concern here. 
  t_ls<-seq(0.1,25,0.1) # time points to check to find the solution of predicted stopping times; one may also consider using uniroot which should provide predictions on a finer scale.
  interim_pred0<-t_ls[which.min(abs(sum_D*fixed_death_p-sapply(t_ls,function(t) {h.tilde(0,lambda,1,eps,R,p,t)})))] ### predicted stoping time at the interim stage under H0
  
  final_pred0<-t_ls[which.min(abs(sum_D-sapply(t_ls,function(t){h.tilde(0,lambda,1,eps,R,p,t)})))] ### predicted stoping time at the final stage under H0
  
  interim_pred1<-t_ls[which.min(abs(sum_D*fixed_death_p-sapply(t_ls,function(t) {h.tilde(0,lambda,theta,eps,R,p,t)})))] ### predicted stoping time at interim stage under H1
  
  final_pred1<-t_ls[which.min(abs(sum_D-sapply(t_ls,function(t){h.tilde(0,lambda,theta,eps,R,p,t)})))]  ### predicted stoping time at final stage under H1
  
  # Predict correlations under H0
  wi_si_cor_pred0<-cor.0(rho1=0,gamma1=0,rho2=0,gamma2=1,lambda=lambda,R=R,p=p,t.star=interim_pred0)
  wf_sf_cor_pred0<-cor.0(rho1=0,gamma1=0,rho2=0,gamma2=1,lambda=lambda,R=R,p=p,t.star=final_pred0)
  si_sf_cor_pred0<-sqrt(I.0(0,0,lambda=lambda,R=R,p=p,t.star=interim_pred0)/I.0(0,0,lambda=lambda,R=R,p=p,t.star=final_pred0))
  wi_wf_cor_pred0<-sqrt(I.0(0,1,lambda=lambda,R=R,p=p,t.star=interim_pred0)/I.0(0,1,lambda=lambda,R=R,p=p,t.star=final_pred0))
  # Follow Theorem 1 in the reference paper 5
  wi_sf_cor_pred0<-wi_si_cor_pred0*si_sf_cor_pred0
  wf_si_cor_pred0<-wi_si_cor_pred0*wi_wf_cor_pred0
  
    # Predict correlations under H1
  wi_si_cor_pred1<-cor.1(rho1=0,gamma1=0,rho2=0,gamma2=1,lambda=lambda,theta=theta,eps=eps,R=R,p=p,t.star=interim_pred1)
  wf_sf_cor_pred1<-cor.1(rho1=0,gamma1=0,rho2=0,gamma2=1,lambda=lambda,theta=theta,eps=eps,R=R,p=p,t.star=final_pred1)
  
  si_sf_cor_pred1<-sqrt(I.1(0,0,lambda=lambda,theta=theta,eps=eps,R=R,p=p,t.star=interim_pred1)/I.1(0,0,lambda=lambda,theta=theta,eps=eps,R=R,p=p,t.star=final_pred1))
  wi_wf_cor_pred1<-sqrt(I.1(0,1,lambda=lambda,theta=theta,eps=eps,R=R,p=p,t.star=interim_pred1)/I.1(0,1,lambda=lambda,theta=theta,eps=eps,R=R,p=p,t.star=final_pred1))
  
  # Follow Theorem 1 in the reference paper 5s  
  wi_sf_cor_pred1<-wi_si_cor_pred1*si_sf_cor_pred1
  wf_si_cor_pred1<-wi_si_cor_pred1*wi_wf_cor_pred1
  
  
  #### For the vector to prepare the correlation matrix in an order of si, wi, sf, wf
  Sigma0_v<-c(wi_si_cor_pred0,si_sf_cor_pred0,wi_sf_cor_pred0,wf_si_cor_pred0,wi_wf_cor_pred0,wf_sf_cor_pred0)
  Sigma1_v<-c(wi_si_cor_pred1,si_sf_cor_pred1,wi_sf_cor_pred1,wf_si_cor_pred1,wi_wf_cor_pred1,wf_sf_cor_pred1)
  
  #Sigma0_v<-1:6
  # obtain the 4x4 symmetric correlation matrix under H0
  Sigma0<-matrix(1, ncol=4,nrow=4)
  Sigma0[upper.tri(Sigma0)]<- Sigma0_v
  Sigma0[lower.tri(Sigma0)]<- t(Sigma0)[lower.tri(t(Sigma0))]
  
  # Sigma0_v<-1:6
  # obtain the 4x4 symmetric correlation matrix under H1
  Sigma1<-matrix(1, ncol=4,nrow=4)
  Sigma1[upper.tri(Sigma1)]<- Sigma1_v
  Sigma1[lower.tri(Sigma1)]<- t(Sigma1)[lower.tri(t(Sigma1))]
  
  ## Obtain he predicted boundaries (exact prediction method)
  ### long z_alpha_vec as the boundary values for GS-MC with K=2 (2 tests) M=2 (2 stages including 1 interm stage): replicative for multiple tests in each stage
  z_alpha_vec_pred<-Maxcombo.bd(Sigma0 = Sigma0,index=c(1,1,2,2),alpha_sp=c(pnorm(z[1],lower.tail = F),alpha))$z_alpha_vec
  ### short z_alpha as the houndary values: unique entry for each stage
  z_alpha_pred<-Maxcombo.bd(Sigma0 = Sigma0,index=c(1,1,2,2),alpha_sp=c(pnorm(z[1],lower.tail = F),alpha))$z_alpha
  
  ### If we only have the final stage, no interim stage, or in other words, if it is not a group sequential design
  z_final_alpha_pred<-Maxcombo.bd(Sigma0 = Sigma0[3:4,3:4],index=c(1,1),alpha_sp=c(alpha))$z_alpha
  
  ## Obtain the predicted mean: 
  ### add the "-" before the calculated E* to be consistent with the later calculations using -WLRT/SLRT other than WLRT/SLRT, to make the test statistics tend to be positive under the alternative. Note that we　do not need to do so if we are not changing the sign of the WLRT/SLRTs. Just make sure that the means and the test statistics are consistent. 
  mu1<-NULL
  mu1[1]<--sample.size_FH(eps,p,b,interim_pred1,omega=max(interim_pred1-R,0),lambda,lambda.trt,0, 0,alpha,beta)$E.star
  mu1[2]<--sample.size_FH(eps,p,b,interim_pred1,omega=max(interim_pred1-R,0),lambda,lambda.trt,rho, gamma,alpha,beta)$E.star
  mu1[3]<--sample.size_FH(eps,p,b,final_pred1,omega=final_pred1-R,lambda,lambda.trt,0, 0,alpha,beta)$E.star
  mu1[4]<--sample.size_FH(eps,p,b,final_pred1,omega=final_pred1-R,lambda,lambda.trt,rho, gamma,alpha,beta)$E.star
  
  ## Obtain the predicted sample sizes
  n_FH_ls<-Maxcombo.sz(Sigma1=Sigma1,mu1=mu1,z_alpha_vec=z_alpha_vec_pred,beta=beta,interim_vec=c(rep(interim_pred1,2),rep(final_pred1,2)),R=R,n_range=size_range,sum_D=sum_D)
  
  n_MCGS[ll]<-n_FH_ls$n
  d_MCGS[ll]<-n_FH_ls$d
  
  n_FH_00_ls[ll]<-size_FH_00$n
  n_FH_01_ls[ll]<-size_FH_01$n
  
  d_FH_00_ls[ll]<-size_FH_00$n_event
  d_FH_01_ls[ll]<-size_FH_01$n_event
  
}
(timecost<- (proc.time()-pt)/len) # average time usage

# The interval when GS-MC requires the least sample size
eps.ls[which(n_MCGS<pmin(n_FH_00_ls,n_FH_01_ls))]

data_n<-c(n_MCGS,n_FH_00_ls,n_FH_01_ls)
data_d<-c(d_MCGS,d_FH_00_ls,d_FH_01_ls)

index<-factor(c(rep("Maxcombo",len),rep("G(0,0)",len),rep("G(0,1)",len)),levels=c("Maxcombo","G(0,0)","G(0,1)"))
data_plot<-data.frame(data_n,data_d,index=index,eps.ls=rep(eps.ls,3))
# Plotting here
(plot1<-ggplot(data=data_plot,aes(x=eps.ls,y=data_n,linetype=index))+
    geom_line(size=0.75)+labs(title=expression(paste("n vs. ",epsilon)),x=expression(paste(epsilon)),y="n")+theme_light()+
    theme(plot.title = element_text(hjust = 0.5))+
    theme(legend.position = "right",legend.title = element_blank()))

#ggsave("size_plot1_new.png",plot1,height=4,width=4)
#ggsave("size_plot1_new.pdf",plot1,height=4,width=4)

(plot2<-ggplot(data=data_plot,aes(x=eps.ls,y=data_d,linetype=index))+
    geom_line(size=0.75)+labs(title=expression(paste("d vs. ", epsilon)),x=expression(paste(epsilon)),y="d")+theme_light()+
    theme(plot.title = element_text(hjust = 0.5))+
    theme(legend.position = "right",legend.title = element_blank()))

#ggsave("size_plot2_new.png",plot2,height=4,width=4)
#ggsave("size_plot2_new.pdf",plot2,height=4,width=4)

```




## Vignette 2: simulation sample for GS-MC (repeat the results in the paper!)
- The codes for all the tables in the [paper](https://arxiv.org/abs/1911.05684 "Hey, direct me to the tedious methodology paper!").
- 48 different cases, with codes in form of S/E in each entry cell, where S stands for the code of the stochastic prediction method, and E stands for the code of the exact prediction method;

- Under the null hypothesis $H_0$: 

 | Violations       | $\theta=0.7$|  $\theta=0.6$|$\theta=0.5$|
|:---------------|:---------------:|:------:|:-----------:|
| no violation    | 1/25| 9/33|17/41|
| violations I&II |3/27|11/35 |19/43|
| violations I&II&III |5/29| 13/37 | 21/45| 
| violations I&II&IV |7/31| 15/39| 23/47 | 

- Under the alternative hypothesis $H_1$: 

 | Violations       | $\theta=0.7$|  $\theta=0.6$|$\theta=0.5$|
|:---------------|:---------------:|:------:|:-----------:|
| no violation    | 2/26| 10/34|18/42|
| violations I&II |4/28| 12/36| 20/44|
| violations I&II&III |6/30 | 14/38 | 22/46|
| violations I&II&IV | 8/32| 16/40| 24/48 |


```{r vignette2}
set.seed(123456) # set an auspicious seed, like your birthday~
date <- "20191218" #optional for the file name

server_num=2 #any value from 1 to 48 to indicate the 48 different combinations of scenario and methods, for details please check the two tables above. 

hypo<-c("alt","null")[server_num%%2+1] #"null" or "alt".
nloops <- c(1000,5000)[server_num%%2+1]  # 2e5 under the null; 5e4 under the alternative in the paper, here use a smaller number for demonstration (1000 for H1, 5000 for H0, to make life easier)
# Rule of thumb: 
#number of simualtions: recommend using >50,000 for type I error, >10,000 for type II error
case<-paste0(c("N","A","NC","AC","NS","AS","NT","AT"),rep(1:6,each=8))[server_num] #length=48

# indicator of using stochastical or non-stochastical prediction method
stoch<-c(T,F)[ifelse(server_num<=24,1,2)]
violation<-c(0,1,2,3)[ceiling(ifelse(server_num%%8==0,8,server_num%%8)/2)] #0 denotes "no violation", 1 denotes "violations on the accrual, censoring"; 2 denotes "in addition to the violations denoted by 1, the event rate is wrong"; 3 denotes "in addition to the violations denoted by 1, the delayed time is wrong"  


fixed_death_p<-0.6 # stop when what proportion of the events have been observed.
p<-1/2 # treatment assignment.
tau<-18 # end of the study, chronological.
R <- 14 # end of the accrual period, chronological.
omega <- (tau-R) # minimum potenital follow-up time, not considering dropouts or other subect specific censoring.
eps <- 2 # awaiting time before the change point.
lambda <- log(2)/6 # event hazard of the control arm.
theta <- c(0.7,0.6,0.5)[ifelse((ceiling(server_num/8))%%3==0,3,(ceiling(server_num/8))%%3)] # hazard ratio after the change point eps under the alternative hypothesis. 
lambda.trt <- lambda*theta # event hazard for the treatment arm under the alternative hypothesis and after the change point eps. 
alpha <- 0.025 # type I error under the control.
beta <- 0.1 # type II error under the control. 

# parameters for WLRT G
rho <- 0
gamma <- 1
# Assume that there is not subject specific censoring
cen_p_ctrl<-1e-6
cen_p_trt<-1e-6
cen_ctrl<--log(1-cen_p_ctrl)/12
cen_trt<--log(1-cen_p_trt)/12

# Special parameters for the Violations 2-3
eps_star=6
lambda_star=log(2)/12

# The title for the record of the output csv file
title <- paste0(date,"_",case)
## boundary calculation using the old method gsDesign for standard G(0,0) log-rank test. 
x <- gsDesign(k=2, test.type=1, timing=0.6, sfu="OF", alpha=alpha, beta=beta,delta=-log(theta))
(z <- x$upper$bound)


# Below is for the boundary and sample size calculation
b <- 30 # number of subintervals for a time unit
# calculat the sample sizes using two different WLRTs, the n needed for the maxcombo should lie between the two values. 
size_FH_upper <- sample.size_FH(eps,p,b,tau,omega,lambda,lambda.trt,0, 0,alpha,beta)
size_FH_lower<-sample.size_FH(eps,p,b,tau,omega,lambda,lambda.trt,0, 1,alpha,beta)
sum_D<-size_FH_upper$sum_D # unit death, should be identical to size_FH_lower$sum_D
# size_seq is the range of the possible sizes using the old method calculation. 
size_range<-c(min(size_FH_lower$n,size_FH_upper$n),max(size_FH_lower$n,size_FH_upper$n))
size_seq<-seq(min(size_FH_lower$n,size_FH_upper$n),max(size_FH_lower$n,size_FH_upper$n),1)
t_ls<-seq(0.1,25,0.1) # a sequence of time to predict the interim and final stage time under the null and alternative hypotheses. 
c=exp(-lambda*(1-theta)*eps) # the c constant for piecewise exponential hazard and survival curves calculation

if(stoch){
  # Under the null, predict the interim and final stage time. 
  interim_pred0<-t_ls[which.min(abs(sum_D*fixed_death_p-sapply(t_ls,function(t) {stoch_pred(eps,p,b,t,omega=max(t-R,0),lambda,1,rho,gamma,R)$sum_D})))]
  
  final_pred0<-t_ls[which.min(abs(sum_D-sapply(t_ls,function(t) {stoch_pred(eps,p,b,t,omega=max(t-R,0),lambda,1,rho,gamma,R)$sum_D})))]
  
  # Under the alternative, predict the inteirm and final stage time.
  interim_pred1<-t_ls[which.min(abs(sum_D*fixed_death_p-sapply(t_ls,function(t) {stoch_pred(eps,p,b,t,omega=max(t-R,0),lambda,theta,rho,gamma,R)$sum_D})))]
  
  final_pred1<-t_ls[which.min(abs(sum_D-sapply(t_ls,function(t) {stoch_pred(eps,p,b,t,omega=max(t-R,0),lambda,theta,rho,gamma,R)$sum_D})))]
  
  
  # Predict the correlation values under the null (indeced by 0) and the alternative (indeced by 1). 
  ## wi: WLRT at interim; wf: WLRT at final; si: SLRT at interim; sf: SLRT at final. 
  ### Under the null
  wi_si_cor_pred0<-stoch_pred.cov(eps,p,b,interim_pred0,omega=max(interim_pred0-R,0),lambda,1,rho,gamma,0,0,R)$cov/sqrt(stoch_pred(eps,p,b,interim_pred0,omega=max(interim_pred0-R,0),lambda,1,rho,gamma,R)$inf*stoch_pred(eps,p,b,interim_pred0,omega=max(interim_pred0-R,0),lambda,1,0,0,R)$inf) 
  
  wf_sf_cor_pred0<-stoch_pred.cov(eps,p,b,final_pred0,omega=max(final_pred0-R,0),lambda,1,rho,gamma,0,0,R)$cov/sqrt(stoch_pred(eps,p,b,final_pred0,omega=max(final_pred0-R,0),lambda,1,rho,gamma,R)$inf*stoch_pred(eps,p,b,final_pred0,omega=max(final_pred0-R,0),lambda,1,0,0,R)$inf) 
  
  si_sf_cor_pred0<-sqrt(stoch_pred(eps,p,b,interim_pred0,omega=max(interim_pred0-R,0),lambda,1,0,0,R)$inf/stoch_pred(eps,p,b,final_pred0,omega=max(final_pred0-R,0),lambda,1,0,0,R)$inf)
  
  wi_wf_cor_pred0<-sqrt(stoch_pred(eps,p,b,interim_pred0,omega=max(interim_pred0-R,0),lambda,1,rho,gamma,R)$inf/stoch_pred(eps,p,b,final_pred0,omega=max(final_pred0-R,0),lambda,1,rho,gamma,R)$inf)
  
  wi_sf_cor_pred0<-wi_si_cor_pred0*si_sf_cor_pred0
  wf_si_cor_pred0<-wi_si_cor_pred0*wi_wf_cor_pred0
  
  ### under the alternative 
  wi_si_cor_pred1<-stoch_pred.cov(eps,p,b,interim_pred1,omega=max(interim_pred1-R,0),lambda,theta,rho,gamma,0,0,R)$cov/sqrt(stoch_pred(eps,p,b,interim_pred1,omega=max(interim_pred1-R,0),lambda,theta,rho,gamma,R)$inf*stoch_pred(eps,p,b,interim_pred1,omega=max(interim_pred1-R,0),lambda,theta,0,0,R)$inf) 
  
  wf_sf_cor_pred1<-stoch_pred.cov(eps,p,b,final_pred1,omega=max(final_pred1-R,0),lambda,theta,rho,gamma,0,0,R)$cov/sqrt(stoch_pred(eps,p,b,final_pred1,omega=max(final_pred1-R,0),lambda,theta,rho,gamma,R)$inf*stoch_pred(eps,p,b,final_pred1,omega=max(final_pred1-R,0),lambda,theta,0,0,R)$inf) 
  
  si_sf_cor_pred1<-sqrt(stoch_pred(eps,p,b,interim_pred1,omega=max(interim_pred1-R,0),lambda,theta,0,0,R)$inf/stoch_pred(eps,p,b,final_pred1,omega=max(final_pred1-R,0),lambda,theta,0,0,R)$inf)
  
  
  wi_wf_cor_pred1<-sqrt(stoch_pred(eps,p,b,interim_pred1,omega=max(interim_pred1-R,0),lambda,theta,rho,gamma,R)$inf/stoch_pred(eps,p,b,final_pred1,omega=max(final_pred1-R,0),lambda,theta,rho,gamma,R)$inf)
  
  wi_sf_cor_pred1<-wi_si_cor_pred1*si_sf_cor_pred1
  wf_si_cor_pred1<-wi_si_cor_pred1*wi_wf_cor_pred1
} else{
  interim_pred0<-t_ls[which.min(abs(sum_D*fixed_death_p-sapply(t_ls,function(t) {h.tilde(0,lambda,1,eps,R,p,t)})))] ### constant ratio between two treatment
  
  final_pred0<-t_ls[which.min(abs(sum_D-sapply(t_ls,function(t){h.tilde(0,lambda,1,eps,R,p,t)})))]### constant ratio between two treatment
  
  interim_pred1<-t_ls[which.min(abs(sum_D*fixed_death_p-sapply(t_ls,function(t) {h.tilde(0,lambda,theta,eps,R,p,t)})))] ### constant ratio between two treatment
  
  final_pred1<-t_ls[which.min(abs(sum_D-sapply(t_ls,function(t){h.tilde(0,lambda,theta,eps,R,p,t)})))]  ### constant ratio between two treatment
  
  wi_si_cor_pred0<-cor.0(rho1=0,gamma1=0,rho2=0,gamma2=1,lambda=lambda,R=R,p=p,t.star=interim_pred0)
  wf_sf_cor_pred0<-cor.0(rho1=0,gamma1=0,rho2=0,gamma2=1,lambda=lambda,R=R,p=p,t.star=final_pred0)
  si_sf_cor_pred0<-sqrt(I.0(0,0,lambda=lambda,R=R,p=p,t.star=interim_pred0)/I.0(0,0,lambda=lambda,R=R,p=p,t.star=final_pred0))
  wi_wf_cor_pred0<-sqrt(I.0(0,1,lambda=lambda,R=R,p=p,t.star=interim_pred0)/I.0(0,1,lambda=lambda,R=R,p=p,t.star=final_pred0))
  wi_sf_cor_pred0<-wi_si_cor_pred0*si_sf_cor_pred0
  wf_si_cor_pred0<-wi_si_cor_pred0*wi_wf_cor_pred0
  
  wi_si_cor_pred1<-cor.1(rho1=0,gamma1=0,rho2=0,gamma2=1,lambda=lambda,theta=theta,eps=eps,R=R,p=p,t.star=interim_pred1)
  wf_sf_cor_pred1<-cor.1(rho1=0,gamma1=0,rho2=0,gamma2=1,lambda=lambda,theta=theta,eps=eps,R=R,p=p,t.star=final_pred1)
  
  si_sf_cor_pred1<-sqrt(I.1(0,0,lambda=lambda,theta=theta,eps=eps,R=R,p=p,t.star=interim_pred1)/I.1(0,0,lambda=lambda,theta=theta,eps=eps,R=R,p=p,t.star=final_pred1))
  wi_wf_cor_pred1<-sqrt(I.1(0,1,lambda=lambda,theta=theta,eps=eps,R=R,p=p,t.star=interim_pred1)/I.1(0,1,lambda=lambda,theta=theta,eps=eps,R=R,p=p,t.star=final_pred1))
  
  wi_sf_cor_pred1<-wi_si_cor_pred1*si_sf_cor_pred1
  wf_si_cor_pred1<-wi_si_cor_pred1*wi_wf_cor_pred1
}

####in an order of si, wi, sf, wf
Sigma0_v<-c(wi_si_cor_pred0,si_sf_cor_pred0,wi_sf_cor_pred0,wf_si_cor_pred0,wi_wf_cor_pred0,wf_sf_cor_pred0)
Sigma1_v<-c(wi_si_cor_pred1,si_sf_cor_pred1,wi_sf_cor_pred1,wf_si_cor_pred1,wi_wf_cor_pred1,wf_sf_cor_pred1)

#Sigma0_v<-1:6
#obtain the 4x4 correlation matrix
Sigma0<-matrix(1, ncol=4,nrow=4)
Sigma0[upper.tri(Sigma0)]<- Sigma0_v
Sigma0[lower.tri(Sigma0)]<- t(Sigma0)[lower.tri(t(Sigma0))]

# Sigma0_v<-1:6
Sigma1<-matrix(1, ncol=4,nrow=4)
Sigma1[upper.tri(Sigma1)]<- Sigma1_v
Sigma1[lower.tri(Sigma1)]<- t(Sigma1)[lower.tri(t(Sigma1))]

## Obtain he predicted boundaries
### long z_alpha_vec
z_alpha_vec_pred<-Maxcombo.bd(Sigma0 = Sigma0,index=c(1,1,2,2),alpha_sp=c(pnorm(z[1],lower.tail = F),alpha))$z_alpha_vec
### short z_alpha
z_alpha_pred<-Maxcombo.bd(Sigma0 = Sigma0,index=c(1,1,2,2),alpha_sp=c(pnorm(z[1],lower.tail = F),alpha))$z_alpha
### For the final stage
z_final_alpha_pred<-Maxcombo.bd(Sigma0 = Sigma0[3:4,3:4],index=c(1,1),alpha_sp=c(alpha))$z_alpha

## Obtain the predicted mean: 
### add the "-" before the calculated E* to be consistent with the later calculations using -WLRT/SLRT other than WLRT/SLRT, to make the test statistics tend to be positive under the alternative. Note need to do so if we are not changing the sign of the WLRT/SLRTs. Just make sure that the means and the test statistics are consistent. 
mu1<-NULL
mu1[1]<--sample.size_FH(eps,p,b,interim_pred1,omega=max(interim_pred1-R,0),lambda,lambda.trt,0, 0,alpha,beta)$E.star
mu1[2]<--sample.size_FH(eps,p,b,interim_pred1,omega=max(interim_pred1-R,0),lambda,lambda.trt,rho, gamma,alpha,beta)$E.star
mu1[3]<--sample.size_FH(eps,p,b,final_pred1,omega=final_pred1-R,lambda,lambda.trt,0, 0,alpha,beta)$E.star
mu1[4]<--sample.size_FH(eps,p,b,final_pred1,omega=final_pred1-R,lambda,lambda.trt,rho, gamma,alpha,beta)$E.star

## Obtain the predicted sample sizes
n_FH_ls<-Maxcombo.sz(Sigma1=Sigma1,mu1=mu1,z_alpha_vec=z_alpha_vec_pred,beta=beta,interim_vec=c(rep(interim_pred1,2),rep(final_pred1,2)),R=R,n_range=size_range,sum_D=sum_D)

n_FH<-n_FH_ls$n
n_event_FH<-n_FH_ls$d
d_fixed<-ceiling(fixed_death_p*n_event_FH) # the number of events needed to pause the study for an interim analysis

# the accrual rate needed
accrual.rt <- n_FH/sum(R)



# Assumption violations: the true accrual rates, censoring rates and event hazard
if(violation==1){
  accrual.rt0<-c(accrual.rt/5,accrual.rt/5*2,accrual.rt/5*3,
                accrual.rt/5*4,accrual.rt/5*6) #total=14*accrual.rt
  R0 <- c(1,1,1,1,10)
  cen_p_ctrl0<-0.1 #yearly censoring probability
  cen_p_trt0<-0.2
  cen_ctrl0<--log(1-cen_p_ctrl0)/12
  cen_trt0<--log(1-cen_p_trt0)/12
  lambda0 <- lambda #log(2)/12
  lambda0.trt <- lambda.trt #log(2)/12*theta
  eps0<-eps
  intervals0<-c(eps0)
  
}else if (violation==2){
  accrual.rt0<-c(accrual.rt/5,accrual.rt/5*2,accrual.rt/5*3,
                 accrual.rt/5*4,accrual.rt/5*6) #total=14*accrual.rt
  R0 <- c(1,1,1,1,10)
  cen_p_ctrl0<-0.1 #yearly censoring probability
  cen_p_trt0<-0.2
  cen_ctrl0<--log(1-cen_p_ctrl0)/12
  cen_trt0<--log(1-cen_p_trt0)/12
  lambda0 <- lambda_star #log(2)/12
  lambda0.trt <- lambda_star*theta #log(2)/12*theta
  eps0<-eps
  intervals0<-c(eps0)
}else if (violation==3){
  accrual.rt0<-c(accrual.rt/5,accrual.rt/5*2,accrual.rt/5*3,
                 accrual.rt/5*4,accrual.rt/5*6) #total=14*accrual.rt
  R0 <- c(1,1,1,1,10)
  cen_p_ctrl0<-0.1 #yearly censoring probability
  cen_p_trt0<-0.2
  cen_ctrl0<--log(1-cen_p_ctrl0)/12
  cen_trt0<--log(1-cen_p_trt0)/12
  lambda0 <- lambda #log(2)/12
  lambda0.trt <- lambda.trt #log(2)/12*theta
  eps0<-eps_star
  intervals0<-c(eps0)
}else{
  accrual.rt0<-accrual.rt
  # true accrual period
  R0<-R
  cen_ctrl0<-cen_ctrl
  cen_trt0<-cen_trt
  lambda0 <- lambda
  lambda0.trt<- lambda.trt
  eps0<-eps
  intervals0<-c(eps0)
}


#############################################
# Assign the predicted correlations according to the known status: either null or alternative hypothesis. 

if(hypo=="null"){
  interim_pred<-interim_pred0
  
  final_pred<-final_pred0
  
  wi_si_cor_pred<-wi_si_cor_pred0
  wf_sf_cor_pred<-wf_sf_cor_pred0
  si_sf_cor_pred<-si_sf_cor_pred0
  wi_wf_cor_pred<-wi_wf_cor_pred0
  wi_sf_cor_pred<-wi_sf_cor_pred0
  wf_si_cor_pred<-wf_si_cor_pred0
}else{
  interim_pred<-interim_pred1
  
  final_pred<-final_pred1
  
  wi_si_cor_pred<-wi_si_cor_pred1
  wf_sf_cor_pred<-wf_sf_cor_pred1
  si_sf_cor_pred<-si_sf_cor_pred1
  wi_wf_cor_pred<-wi_wf_cor_pred1
  wi_sf_cor_pred<-wi_sf_cor_pred1
  wf_si_cor_pred<-wf_si_cor_pred1
}

# Start of the simulations and the data-driven boundary estimation. 
z_alpha_est<-matrix(NA,nrow=nloops,ncol=2)
z_final_est=z1_est=z2_est=NULL
maxcombo.final.dec.onestep<-maxcombo.interim<-maxcombo.final<-maxcombo.interim.dec<-maxcombo.final.dec<-NULL
# Definition for the estimated correlations below. 
wlrt.interim <- wlrt.final <- NULL
slrt.interim <- slrt.final <- NULL
final_time <- interim_time <- NULL
wi_sf_cor_est<-wi_si_cor_est<-si_sf_cor_est<-NULL
wf_si_cor_est<-wf_sf_cor_est<-wi_wf_cor_est<-NULL
# wi_wf_cor_est2<-NULL
Sigma_v<-matrix(NA,nrow=nloops,ncol=6)
#pb <- txtProgressBar(min = 0, max = nloops, style = 3)
for(l in  1:nloops){
  set.seed(l)
  {if(hypo=="null") data_temp<- nphsim(nsim=1,lambdaC=lambda0, lambdaE = lambda0, ssC=ceiling(n_FH*(1-p)), intervals = intervals0,ssE=ceiling(n_FH*p), gamma=accrual.rt0, R=R0, eta=cen_ctrl0,etaE=cen_trt0, fixEnrollTime = TRUE)$simd # fixed enrollment time, adjust the accrual time
    else if (hypo=="alt") data_temp<- nphsim(nsim=1,lambdaC=lambda0, lambdaE = c(lambda0,lambda0.trt), ssC=ceiling(n_FH*(1-p)), intervals = intervals0,ssE=ceiling(n_FH*p), gamma=accrual.rt0, R=R0, eta=cen_ctrl0,etaE=cen_trt0, fixEnrollTime = TRUE)$simd # fixed enrollment time, adjust the accrual time
  }
  data_interim<-data.trim.d(d_fixed,data_temp)
  data_final<-data.trim.d(n_event_FH,data_temp)
  
  interim_time[l]<-data_interim[[2]]
  final_time[l]<-data_final[[2]]
  
  data_interim=data_interim[[1]]
  data_final=data_final[[1]]
  
  wlrt.interim[l]<- tryCatch(FH.test(survival=data_interim$survival,delta=data_interim$delta,trt=data_interim$trt,rho=rho,gamma=gamma)$Z,error=function(e) NA)
  slrt.interim[l]<- tryCatch(FH.test(survival=data_interim$survival,delta=data_interim$delta,trt=data_interim$trt,rho=0,gamma=0)$Z,error=function(e) NA)
  w2<-function(...){survKM_minus(...)^rho*(1-survKM_minus(...))^gamma}
  wi_si_cor_est[l]<- WLR.test.cor(survival=data_interim$survival,delta=data_interim$delta,trt=data_interim$trt,w2=w2)
  
  wlrt.final[l]<- tryCatch(FH.test(survival=data_final$survival,delta=data_final$delta,trt=data_final$trt,rho=rho,gamma=gamma)$Z,error=function(e) NA)
  slrt.final[l]<- tryCatch(FH.test(survival=data_final$survival,delta=data_final$delta,trt=data_final$trt,rho=0,gamma=0)$Z,error=function(e) NA)
  wf_sf_cor_est[l]<- WLR.test.cor(survival=data_final$survival,delta=data_final$delta,trt=data_final$trt,w2=w2)
  wi_wf_cor_est[l]<-sqrt(I_t(data_ref=data_final,data_check= data_interim,rho=rho,gamma=gamma)/I_t(data_ref=data_final,data_check = data_final,rho=rho,gamma=gamma)  )
 
  si_sf_cor_est[l]<-sqrt(fixed_death_p) 
 
  wi_sf_cor_est[l]<-wi_si_cor_est[l]*si_sf_cor_est[l]
  
  wf_si_cor_est[l]<-wi_si_cor_est[l]*wi_wf_cor_est[l]
  
  # Data-driven alpha spending: 
  Sigma_v[l,]<-c(wi_si_cor_est[l],si_sf_cor_est[l],wi_sf_cor_est[l],wf_si_cor_est[l],wi_wf_cor_est[l],wf_sf_cor_est[l])
  # Sigma0_v<-1:6
  Sigma<-matrix(1, ncol=4,nrow=4)
  Sigma[upper.tri(Sigma)]<- Sigma_v[l,]
  Sigma[lower.tri(Sigma)]<- t(Sigma)[lower.tri(t(Sigma))]
  # O'Brien and Fleming test (OBF)
  z_alpha_est[l,]<-Maxcombo.bd(Sigma0 = Sigma,index=c(1,1,2,2),alpha_sp=c(pnorm(z[1],lower.tail = F),alpha))$z_alpha
  
  z1_est[l]<- z_alpha_est[l,1]
  z2_est[l]<- z_alpha_est[l,2]
  
  # Invert the sign to make most test statistics positive under the alternative hypothesis
  maxcombo.interim<-max(-wlrt.interim[l],-slrt.interim[l])
  maxcombo.final<-max(-wlrt.final[l],-slrt.final[l])
  maxcombo.interim.dec[l]<-(maxcombo.interim>z1_est[l])
  maxcombo.final.dec[l]<-(maxcombo.final>z2_est[l])
  
  # For only one-stage case
  z_final_est[l]<-Maxcombo.bd(Sigma0 = Sigma[3:4,3:4],index=c(1,1),alpha_sp=c(alpha))$z_alpha
  maxcombo.final.dec.onestep[l]<-(maxcombo.final>z_final_est[l])
  
  #setTxtProgressBar(pb, l)
}
#close(pb)

sum(!complete.cases(wlrt.interim)) 
sum(!complete.cases(wlrt.final)) 
sum(!complete.cases(slrt.interim)) 
sum(!complete.cases(slrt.final)) 

maxcombo.interim<-pmin(wlrt.interim,slrt.interim) # need to take the inverse sign
maxcombo.final<-pmin(wlrt.final,slrt.final)

(wi_si_cor<-cor(wlrt.interim,slrt.interim,use="complete.obs")) # wlrt.interim,slrt.interim
mean(wi_si_cor_est)
wi_si_cor_pred

(wf_sf_cor<-cor(wlrt.final,slrt.final,use="complete.obs"))     # wlrt.final,slrt.final
mean(wf_sf_cor_est)
wf_sf_cor_pred

(si_sf_cor<-cor(slrt.interim,slrt.final,use="complete.obs"))   # slrt.interim,slrt.final
mean(si_sf_cor_est)
si_sf_cor_pred

(wi_wf_cor<-cor(wlrt.interim,wlrt.final,use="complete.obs"))   # wlrt.iterim,wlrt.final
mean(wi_wf_cor_est)
#mean(wi_wf_cor_est2)
wi_wf_cor_pred

(wi_sf_cor<-cor(wlrt.interim,slrt.final,use="complete.obs"))  # wlrt.interim and slrt.final
mean(wi_sf_cor_est)
wi_sf_cor_pred

(wf_si_cor<-cor(wlrt.final,slrt.interim,use="complete.obs"))   # wlrt.final and slrt.interim
mean(wf_si_cor_est)
wf_si_cor_pred

# mi: maxcombo at interim; mf: maxcombo at final
(mi_mf_cor<-cor(maxcombo.interim,maxcombo.final,use="complete.obs")) 

## Good
wi_si_cor*si_sf_cor
wi_sf_cor
wi_sf_cor_pred


## Good
wi_si_cor*wi_wf_cor
wf_si_cor
wf_si_cor_pred



(wlrt1<-mean(-wlrt.final>qnorm(0.025,lower.tail = F),na.rm=T)) 
(wlrt2<-1-mean((-wlrt.interim)<z[1]&(-wlrt.final)<z[2],na.rm=T)) 
(wlrt2_interim<-1-mean(-wlrt.interim<z[1],na.rm=T)) 


(slrt1<-mean(-slrt.final>qnorm(0.025,lower.tail = F),na.rm=T))
(slrt2<-1-mean((-slrt.interim)<z[1]&(-slrt.final)<z[2],na.rm=T)) 
(slrt2_interim<-1-mean(-slrt.interim<z[1],na.rm=T) )


(mc1<-mean(-maxcombo.final>qnorm(0.025,lower.tail = F),na.rm=T))
(mc2<-1-mean((-maxcombo.interim)<z[1]&(-maxcombo.final)<z[2],na.rm=T)) 
(mc2_interim<-1-mean(-maxcombo.interim<z[1],na.rm=T)) 

interim_pred
mean(interim_time,na.rm = T)
final_pred
mean(final_time,na.rm = T)


(mc1_pred<-mean(-maxcombo.final>z_final_alpha_pred,na.rm=T) )
(mc2_pred<-1-mean((-maxcombo.interim)<z_alpha_pred[1]&(-maxcombo.final)<z_alpha_pred[2],na.rm=T)) 
(mc2_interim_pred<-1-mean(-maxcombo.interim<z_alpha_pred[1],na.rm=T) )


(mc1_est<-mean(maxcombo.final.dec.onestep))
(mc2_est<-mean(maxcombo.interim.dec)+mean(maxcombo.final.dec)-mean(maxcombo.interim.dec*maxcombo.final.dec))
(mc2_interim_est<-mean(maxcombo.interim.dec))

options(digits=5)

out<-rbind(size_wlrt=size_FH_lower[[1]],size_slrt=size_FH_upper[[1]],n_FH,n_event_FH,sum_D,old_z1=z[1],old_z2=z[2],old_final_z=qnorm(p=0.975,lower.tail =T),
           z_alpha_pred1=z_alpha_pred[1],z_alpha_pred2=z_alpha_pred[2],z_final_alpha_pred,z1_est_mean=mean(z1_est),z2_est_mean=mean(z2_est),z_final_est=mean(z_final_est),z1_est_sd=sd(z1_est),z2_est_sd=sd(z2_est),z_final_sd=sd(z_final_est),wi_si_cor,wi_si_cor_est=mean(wi_si_cor_est),wi_si_cor_pred,
           wf_sf_cor,wf_sf_cor_est=mean(wf_sf_cor_est),
           wf_sf_cor_pred,si_sf_cor,si_sf_cor_est=mean(si_sf_cor_est),
           si_sf_cor_pred,wi_wf_cor,wi_wf_cor_est=mean(wi_wf_cor_est),
           wi_wf_cor_pred,wi_sf_cor,wi_sf_cor_est=mean(wi_sf_cor_est),
           wi_sf_cor_pred,wf_si_cor,wf_si_cor_est=mean(wf_si_cor_est),
           wf_si_cor_pred,mi_mf_cor,interim_pred,interim_time=mean(interim_time,na.rm = T),
           final_pred,final_time=mean(final_time,na.rm = T),
           wlrt1,wlrt2,wlrt2_interim,
           slrt1,slrt2,slrt2_interim,
           mc1,mc2,mc2_interim,
           mc1_pred,mc2_pred,mc2_interim_pred,
           mc1_est,mc2_est,mc2_interim_est)
out
title
#write.csv(out,paste0(title,".csv"))

```







## References
1. Lakatos, E. (1988). Sample sizes based on the log-rank statistic in complex clinical trials. Biometrics, 229-241.

2. Hasegawa, T. (2014). Sample size determination for the weighted log‐rank test with the Fleming–Harrington class of weights in cancer vaccine studies. Pharmaceutical statistics, 13(2), 128-135.

3. Hasegawa, T. (2016). Group sequential monitoring based on the weighted log‐rank test statistic with the Fleming–Harrington class of weights in cancer vaccine studies. Pharmaceutical statistics, 15(5), 412-419.

4. Luo, X., Mao, X., Chen, X., Qiu, J., Bai, S., & Quan, H. (2019). Design and monitoring of survival trials in complex scenarios. Statistics in medicine, 38(2), 192-209.

5. Wang, L., Luo, X., & Zheng, C. (2019). A Simulation-free Group Sequential Design with Max-combo Tests in the Presence of Non-proportional Hazards. arXiv preprint arXiv:1911.05684.

Dependent R packages:
[1]: https://github.com/lilywang1988/IAfrac 

[2]: https://github.com/keaven/nphsim

[3]: https://cran.r-project.org/web/packages/mvtnorm/index.html
